---
title: "p8105_hw5_lg3450"
output: github_document
---

first load all necessary libraries

```{r message=FALSE}
library(tidyverse)
library(broom)
library(scales)   
library(dplyr)
```

# Problem 1

**Suppose you put n people in a room, and want to know the probability that at least two people share a birthday. For simplicity, we’ll assume there are no leap years (i.e. there are only 365 days) and that birthdays are uniformly distributed over the year (which is actually not the case).**

-   Q1.1: Write a function that, for a fixed group size, randomly draws “birthdays” for each person; checks whether there are duplicate birthdays in the group; and returns TRUE or FALSE based on the result.

```{r}
check_birthday = function(n) {
  #randomly draw #n birthdays from 1 to 365, allow for same bthd
  birthdays = sample(1:365, size = n, replace = TRUE)
  #logic: if length < n, duplication appears
  repeated_bthd = length(unique(birthdays)) < n
  repeated_bthd}
```

-   Q1.2: Next, run this function 10000 times for each group size between 2 and 50. For each group size, compute the probability that at least two people in the group will share a birthday by averaging across the 10000 simulation runs.

```{r}
set.seed(1)

prob_results =
  expand_grid(
    group_size = 2:50,
    iter = 1:10000) |>
  mutate(
    repeated_bthd = map_dbl(group_size, check_birthday)) |>
  group_by(group_size) |>
  summarize(
    prob_shared = mean(repeated_bthd))

prob_results
```

-   Q1.3: Make a plot showing the probability as a function of group size, and comment on your results.

```{r}
prob_results |>
  ggplot(aes(x = group_size, y = prob_shared)) +
  geom_point() +
  geom_path() +
  labs(
    title = "Estimated Probability of a Shared Birthday",
    x = "Group Size",
    y = "Probability of at least one shared birthday"
  ) +
  scale_y_continuous(labels = percent)
```

The plot illustrates a clear non-linear acceleration in probability, rather than a steady, linear increase. This simulation confirms the "birthday paradox," showing the 50% probability threshold is crossed with a group size of just 23. Following this point, the probability rapidly converges toward certainty, reaching 97% by a group size of 50.

# Problem2
-   Q2.1: Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of mu on the x axis. Describe the association between effect size and power.

```{r}
set.seed(1)

n = 30
sigma = 5
mu_values = 0:6
n_sims = 5000
alpha = 0.05

run_one = function(mu) {
 x = rnorm(n, mean = mu, sd = sigma)
 t_out = t.test(x, mu = 0) |> tidy()
 tibble(
  true_mu = mu,
  mu_hat = as.numeric(t_out$estimate),
  p_value = t_out$p.value)}

#run all simulations for mu = 0,1,2,3,4,5,6
sim_results =
 expand_grid(true_mu = mu_values, iter = 1:n_sims) |>
 mutate(res = map(true_mu, run_one)) |>
 select(res) |>
 unnest(res) |>
 mutate(reject = p_value < alpha)

power_df =
 sim_results |>
 group_by(true_mu) |>
 summarize(power = mean(reject))

power_plot =
 power_df |>
 ggplot(aes(x = true_mu, y = power)) +
 geom_point(size = 2) +
 geom_line() +
 scale_y_continuous(labels = percent, limits = c(0, 1)) +
 labs(
  title = "Power of One-Sample t-test vs. Effect Size",
  x = "True Value of μ (Effect Size)",
  y = "Power (Proportion of Null Rejections)") +
 theme_minimal()

power_plot
```

Based on the plot, the association between and power is strong, positive, and non-linear. When the effect size is 0 (namely the null hypothesis is true), the power is low, at approximately 5%. As the effect size increases, the power (the probability of correctly detecting the effect) rises rapidly. The power is low for a small effect size, but approaches 100% for large effect sizes. In short, the larger the true effect, the more likely the test is to detect it.

-   Q2.2: Make a plot showing the average estimate of mu on the y axis and the true value of mu on the x axis. 
Make a second plot (or overlay on the first) the average estimate of mu only in samples for which the null was rejected on the y axis and the true value of mu on the x axis. 
Is the sample average of mu across tests for which the null is rejected approximately equal to the true value of mu? 
Why or why not?

```{r}
est_df =
 sim_results |>
 group_by(true_mu) |>
 summarize(
  avg_all   = mean(mu_hat),
  avg_rejected = mean(mu_hat[reject], na.rm = TRUE)) |>
 pivot_longer(
  cols = c(avg_all, avg_rejected),
  names_to = "type",
  values_to = "avg_mu_hat")

estimate_plot =
 est_df |>
 ggplot(aes(x = true_mu, y = avg_mu_hat, color = type)) +
 geom_point(size = 2) +
 geom_line() +
 geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
 scale_color_discrete(
  labels = c(
   avg_all = "Average (All Samples)",
   avg_rejected = "Average (Rejected Only)")) +
 labs(
  title = "Average μ̂ vs True μ",
  x = "True μ",
  y = "Average μ̂",
  color = "Estimate Type") +
 theme_minimal()

estimate_plot
```

No. The sample average of $\hat{\mu}$ among tests that reject the null is not equal to the true value of $\mu$ and it is consistently biased upward.

This happens because of selection bias. When we keep only “significant” results (p < 0.05), we disproportionately include simulations where $\hat{\mu}$ happened to be larger than the truth (far from 0), and exclude those closer to 0 (often non-significant). As a result, underestimates are removed while overestimates remain, causing the average of the rejected samples to be inflated.

# Problem3

**The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. You can read their accompanying article here.**

-   Q3.1: Describe the raw data. Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”).

```{r message=FALSE}
homicides = read_csv("../homicide-data.csv")
```

```{r}
unsolved_types = c("Open/No arrest", "Closed without arrest")

city_summary = homicides |>
  mutate(city_state = paste(city, state, sep = ", ")) |>
  mutate(unsolved = disposition %in% unsolved_types) |>
  group_by(city_state) |>
  summarize(
    total_homicides = n(),                      
    unsolved_homicides = sum(unsolved, na.rm = TRUE)) |>
  arrange(desc(total_homicides))

print(head(city_summary, 10))
```

The dataset contains `r nrow(homicides)` individual homicide records and `r ncol(homicides)` variables describing each case.

Key geographic variables include `city` and `state`, which identify the location. In total, the dataset spans `r length(unique(homicides$city))` distinct cities. For more precise spatial analysis, the dataset also provides the specific `lat` (latitude) and `lon` (longitude) for each incident.

The dataset also includes victim demographic variables, such as their `victim_first` and `victim_last` name. For analysis, the most relevant demographic fields are `victim_race` and `victim_sex`, both of which are categorical variables, and `victim_age`.

The disposition variable has three distinct levels. Running `r table(homicides$disposition)` counts the total number of cases for each of these levels, which are: 'Closed by arrest,' 'Open/No arrest,' and 'Closed without arrest.'

-   Q3.2: For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.

```{r}
final_baltimore_estimates =
  homicides |>
  filter(city == "Baltimore", state == "MD") |>
  summarize(
    total_homicides = n(),
    unsolved_homicides = sum(disposition %in% unsolved_types)) |>
  mutate(
    prop_test = list(
      prop.test(
        x = unsolved_homicides,
        n = total_homicides))) |>
  mutate(tidy_res = map(prop_test, broom::tidy)) |>
  select(tidy_res) |>
  unnest(tidy_res) |>
  select(
    estimate_proportion = estimate,
    conf_low = conf.low,
    conf_high = conf.high)

print(final_baltimore_estimates)
```

-   Q3.3: Now run prop.test for each of the cities in your dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.

```{r warning=FALSE}
final_estimates =
  city_summary |>
  mutate(
    tidy_results = purrr::map2(
      .x = unsolved_homicides, 
      .y = total_homicides,
      .f = ~ tidy(prop.test(x = .x, n = .y)))) |>
  unnest(tidy_results) |>
  select(
    city_state,
    total_homicides,
    unsolved_homicides,
    estimate_proportion = estimate,
    conf_low = conf.low,
    conf_high = conf.high
  ) |>
  arrange(desc(estimate_proportion))

print(final_estimates)
```

-   Q3.4: Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.
```{r}
homicide_ci_plot =
  final_estimates |>
  ggplot(aes(x = estimate_proportion, 
             y = fct_reorder(city_state, estimate_proportion))) +
  geom_errorbar(
    aes(xmin = conf_low, xmax = conf_high),
    width = 0.2, 
    color = "darkgrey", 
    alpha = 0.8) +
  geom_point(
    aes(size = total_homicides), 
    color = "darkblue", 
    alpha = 0.6) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(
    title = "Unsolved Homicide Rate by City",
    x = "Estimated Proportion of Unsolved Homicides",
    y = "City",
    size = "Total Homicides") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    panel.grid.major.y = element_blank())

ggsave("homicide_city_CIs_plot.png", homicide_ci_plot, width = 45, height = 45, dpi = 300)

print(homicide_ci_plot)
```

